Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2022-11-05 paper/manuscript.Rmd
Add abstract and significance statements

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
  bookdown::word_document2:
    fontsize: 12pt
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
    fontsize: 12pt
    header-includes:
       - \usepackage{natbib}
       - \usepackage{float}
       - \usepackage{multirow}
       - \usepackage{booktabs}
       - \usepackage{makecell}
       - \usepackage{setspace}
       - \doublespacing
       - \usepackage{lineno}
       - \linenumbers
    geometry: margin=2cm
    number_sections: true
bibliography: references.bib
---

@@

```{r, results = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(viridis)
library(kableExtra)
library(sf)
mypath <- rprojroot::find_package_root_file
source(mypath("R/variable_shortcut.R")) # Define your custom code as a bunch of functions.
source(mypath("R/misc.R"))
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

@@

```{r}
tar_load(full_data2)
nb_st <- unique(full_data2$station) %>% length
rm(full_data2)
```

# Abstract (250 words max)

# Signifiance (120 words)

# Introduction (564 words)

Global diversity is reported to decline at alarming rates [@]. At local scale,
recent meta-analysis reported a huge variation in species richness temporal
trends across various taxa and ecosystem types[@vellend_global_2013;
@dornelas_assemblage_2014; @blowes_geography_2019], with no net variations in
species richness[@dornelas_assemblage_2014; @blowes_geography_2019].
Futhermore, recent studies founs that temporal changes in species richness are
unrelated to changes in species composition[@dornelas_assemblage_2014;
@blowes_geography_2019], rising the recognition of the multidimensional nature
of biodiversity changes[@hillebrand_biodiversity_2018; @blowes_local_nodate]
and the need for further investigations on how biodiversity changes over time.

Species interactions among species, described as ecological networks, drive for
a large part community composition, species coexistence, and ecosystem
functioning[@chesson_mechanisms_2000; @thebault_food-web_2003;
@oconnor_general_2017]. The structure of ecological networks can trigger or
impede perturbation propagations, i.e secondary extinctions[@paine_note_1969;
@dunne_network_2002; @srinivasan_response_2007;
@stouffer_compartmentalization_2011]. Higher connectance, which measures the
degree of generalism in a network, is predicted to increase robustness to
secondary extinctions[@dunne_network_2002; @sole_complexity_2001] and studies
found that higher trophic level was associated to a lower temporal variability
of community biomass in theoretical models and empirical
data[@shanafelt_stability_2018; @danet_species_2021]. Then, accounting for how
the structure of ecological networks changes over time according to other
biodiversity facets is crucial.

Th availability of ecological networks timeseries is
scarce[@kortsch_disentangling_2021], such as predictions about how the strucure
of ecological networks change according to changes in species richness come from
spatial comparison. Spatially, connectance has been shown to be negatively
related to species richness in small (i.e. below 20 species) empirical food-webs
[reviewed in @dunne_network_2006], which reflects constraints on the number of
realizable species interactions or an increasing proportion of specialists in
more rich foodwebs. Connectance has been shown to scale positively or negatively
with community biomass in theoretical food-webs[@thebault_food-web_2003;
@poisot_trophic_2013] while the average trophic level was previously shown to
scale negatively with species richness[@danet_species_2021].

@@

The increasing availability ecological temporal series and the development of
methods to reconstruct food-webs[@bonnaffe_comparison_2021;
@gravel_inferring_2013; @danet_species_2021] makes possible to overcome
those research gaps. We used a standardized monitoring of fish riverine 
communities on the period 1995-2018 over France, documenting the individual size
of fishes in 400 sites during ten years or more. We restructed around 4000
food-webs based on allometric relationships and food diet data, and then
computed connectance, and average trophic level as well as community biomass and
species richness.

Our primary objective was to assess how the local structure of foodwebs changes
over time according to simultaneous changes in community biomass and species
richness over time. Our second objective was to assess if the relationships
between changes in food-web structure, community biomass and species richness,
are the same over space and over time, i.e. is the space-for-time substitution
valid?  To do so, we first assessed the temporal trends in food-web structure.
We then built structural equation model to assess the relationships among the
temporal trends of the different facets of food-web structure.

@@

Using a hierarchical model to estimate conjointly the temporal trends of  
foodweb strucure at each site, we found weak evidence of a global decrease in
community biomass (CI80%:`r gbm` by decade; median [Credible Interval]) but of a
global increase in species richness (CI80%: `r gr` by decade) across the `r nb_st` sites. 
We did not find evidence of a global trends in connectance and average trophic
level (CI80%: `r gct` and `r gt` respectively). The hierarchical model captured
well the variance associated to the spatial structure of the data with
conditional $R^2$ going from  `r r2ic$log_bm_std` (mean [CI95%]) for community
biomass to `r r2ic$log_rich_std` for species richness (Table SXX).

```{r}
sem_cap <- "(a) Structural equation models linking the temporal trends of the food-web
structure. The path widths are proportional to the standardized slope coefficients. The standardized slope coefficients and their associated 95%
  confidence intervals are displayed along the path. (b to f) Direct bivariate
relationships derived from the structural equation model. The distribution of the
site temporal trends are displayed in Fig. SXX."
```

@@

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r rdtps$ct$bm`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that decreased
in biomass over time lose highly connected trophic species over time, and
reversely that communities gaining biomass over time also gain highly connected
trophic species over time. We found a negative relationship between the temporal
trends of connectance and the ones of species richness
($r_\delta = `r rdtps$ct$r`$, Fig. \@ref(fig:sem)d). We further found the
species richness had an indirect positive effect on connectance, through
community biomass ($r_\delta = `r ridtps$ct$r`$). It resulted in a relatively
weak total negative effect of species richness on connectance
($r_\delta = `r rttps$ct$r`$).

@@

Most temporal trends of for species richness and community biomasss were
classified as stable. Roughly the same number of stations displayed temporal
decrease and increase for species richness, and twice more stations displayed
decrease in community biomass than increase. This result is in accordance with
previous meta analysis on species richness trends of vertebrate in marine,
freshwater and terrestrial systems, which show no change in species richness in
average [@dornelas_assemblage_2014; @blowes_geography_2019] but an increase in
total abundance [@van_klink_meta-analysis_2020].

## Species richness and community biomass are positively linked in space and time

Spatially and temporally, species richness scaled positively with community
biomass. Spatially, this result is in accordance with previous studies in stream
communities [@woods_testing_2020; @danet_species_nodate] but also in grasslands,
salt marches and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018; but see @grace_integrative_2016].
Therefore, our results confirm that Biodiversity-Ecosystem functioning
prediction also hold in the temporal dynamics of food-webs. The positive
relationship between species richness and total biomass might be due to positive
complementarity in resource use [@carey_determining_2011], which is held by
niche differentiation [@loreau_biodiversity_1998; @loreau_partitioning_2001] or
to the higher probability of getting high biomass species with higher species
richness [@loreau_biodiversity_2001].

## Food-web structure changes according to the temporal trends of community biomass and species richness

While we found a strong relationship between the temporal trends of community
biomass and the one of species richness, our results suggest that temporal
trends in community biomass and species richness are associated with very
different temporal trends on the connectance and average trophic level. Indeed,
temporal trends of connectance and average trophic level were positively
associated to the temporal trends of community biomass while we found
that species richness temporal trends was negatively and no
significantly related to connectance and average trophic level respectively. All
together, it suggests that temporal trends of community biomass and species
richness are related to different assembly and disassembly processes in
food-webs.

The temporal trends of connectance were positively linked to that of community
biomass, meaning that when connectance increased through time when community
biomass increased through time in a given site. Reversely, connectance decreased
in food-webs when community biomass decreased through time. Indeed, food-webs
have a highly nested structure, in a way that feeding trophic links of species
at lower trophic level are a subset of the ones of species found at higher
trophic level [@kondoh_food_2010; @nordstrom_nestedness_2015]. It is because
feeding interactions are largely driven by body size, consumers being larger
than their preys [@brose_allometric_2006; @woodward_body_2005]. We also found
that community biomass losses through time were associated to a temporal
decrease of average trophic level and that community biomass temporal gains were
associated to temporal increase of average trophic level. Together with
connectance, it suggests that community biomass losses and gains are associated
to losses and gains in large bodied piscivores, i.e. biomass at the higher
trophic level. In the same line, we found that the temporal trends of community
biomass were positively associated to the temporal trends of both number and
proportion of piscivorous fish species present in the communities (Fig.
SXX). The close association between the temporal trends of community biomass
from one side, and connectance, average trophic level from the other side,
strongly suggest that temporal trends in community biomass in natural freshwater
fish communities are mainly associated to variations in the biomass of the top
trophic level.

In contrast, we found that temporal trends of connectance was weakly negatively
related to the temporal trends of species richness, in contrasts with the strong
negative relationships that we found in the present study and from previous
results over spatial gradient [@dunne_network_2006; @danet_species_nodate].
Furthermore, we found that temporal trends of the proportion of piscivorous
species was negatively associated to the temporal trends of species richness
(Table SXX), in accordance with previous meta-analysis over spatial gradient
[@turney_pyramids_2016] and mesocosm experiment [@downing_ecosystem_2002]. It
resonates with the energetic paradigm of food-webs which proposes that the
presence of species at higher trophic level is constrained by energy loss along
the food chain [@lindeman_trophic-dynamic_1942]. It suggests that the temporal
increase of species richness found in our dataset are associated to a decrease
of connectance because species richness increases are biased toward
non-piscivorous species which have less feeding links than piscivorous species.
We found in the same time that the temporal trends species richness had no
direct link with the temporal trends of the average trophic level, the latter
being weighted by the biomass of each node. An explanation can be that when
species richness increases or decreases through time without community biomass
trends, the associated decrease and respectively increase in proportion of
piscivorous species leads to a redistribution of biomass among remaining
piscivorous species or among the non-piscivorous species.

The temporal trends of species richness had an indirect positive effect on the
temporal trends of both connectance and average trophic level through the
positive relationship between the temporal trends of species richness and
community biomass. The temporal trends of species richness had then an
total effect on the temporal trends of connectance and average trophic level
which was twice lower in magnitude than the magnitude of the total effect of the
temporal trends of community biomass. It suggests that the temporal assembly and
disassembly of food-webs are more sensible to variations in community biomass
than in species richness, in other words that food-web structure is more robust
to species richness change than biomass changes. This result is in line with
evidences that food-web containing generalist predators can rewire in response
to changes in the availability of prey species [@bartley_food_2019]. This result
also resonates with the energy paradigm of food-webs which proposes that
food-webs are primarily constrained by fluxes of matter and energy
[@lindeman_trophic-dynamic_1942], so community biomass and not by species
richness.

---
# - Remove the following paragraph:  
# As studies reported community biomass or species richness temporal trends
# [@hallmann_more_2017; @ceballos_biological_2017; @dornelas_assemblage_2014;
# @blowes_geography_2019], we document that temporal trends of species richness
# and community biomass might have different consequences on the structure of the
# community. In particular, the decrease of the average trophic level associated
# to the loss of community biomass could have cascading effect on ecosystem
# functioning as the highest predators are predicted to increase productivity and
# stabilise communities [@wang_biodiversity_2018; @shanafelt_stability_2018;
# @danet_species_nodate]. In contrasts, the loss of species richness might be
# linked to loss of redundancy in the food-web, highlighting that those
# communities may become more at risk regarding future perturbations
# [@yachi_biodiversity_1999]. Interestingly, the hypothesis that temporal trends
# in community biomass is related to higher trophic level in our system has a
# strong reasoning with the effect of human fisheries in marine systems which
# affect primarily fish species at top trophic level [@sprules_surfing_2016].
# Overall, our study highlights that monitoring several facets of biodiversity is
# essential to infer how ecological communities are reshaped by loss and gain of
# biomass and species richness.
---

## Space-for-time substitution does not always hold

In time and space, we found that connectance was positively linked to community
biomass. This result is also in accordance with theoretical work showing that
higher food-web connectance can result in higher community biomass
[@thebault_food-web_2003], or lower [@poisot_trophic_2013], highlighting that
the relation between connectance and community biomass may be dependant on how
food-web structure change with community biomass. Likewise, we found that
average trophic level had a positive relationship with community biomass in
space and time, in accordance with a theoretical study predicting an increase of
community biomass subsequent to the presence of higher predators
[@wang_biodiversity_2018]. Finally, we found a positive relationship between
community biomass and species both in space and time, however the relationship
in time was stronger.

Surprisingly, we found that the temporal trends of connectance were weakly
related to the one of species richness as found spatially in most empirical
food-webs [@dunne_network_2006] and also here. We further found a weak negative
relationship between the temporal trends of average trophic level and the one of
species richness which we found a twice stronger but positive relationship
between median average trophic level and median species richness over the
spatial gradient. An explanation might be that environmental gradient, in
resources in hydrologic contraints might be responsible for the strong negative
effect of species richness on both connectance and average trophic level over
spatial gradient. However, we controlled our spatial analysis by the Biological
Demand in Oxygen, water temperature and the stream gradient. This result is also
in line with other studies which found mismatch between spatial and temporal
patterns [@adler_contrasting_2007; @johnson_testing_2008]. This discrepancy in
space and time might be due to the crucial assumption that variations in space
reflect differences in assembly stages of ecological communities and then
mirrors their temporal assembly/disassembly trajectories. Temporal variation of
species richness might relate on transitory dynamics while spatial gradient
reflect aggregated dynamics over longer time scale including a far number of
processes including high species turnover along spatial gradient.

@@

Our study suggests that species richness and community biomass temporal trends
in natural fish communities relate on different ecological processes, biomass
temporal dynamics being related to variations of food-web structure at the
higher trophic levels while variations of species richness is weakly related to
change in food-web structure. Our result also suggests that space for time
substitution may be relevant to predict the effects of community biomass losses
or gains on food-web structure. However, our findings suggest that
space-for-time substitution might not work for species richness. Our study calls
for further investigations to support the generality of assembly rules in
food-webs. They are crucial to better understand the assembly and disassembly of
food-web and then predict its impact on ecosystem functioning.

@@

We used a monitoring of freshwater fishes conducted all over France and
conducted by the Office Français de la Biodiversité (OFB, previously known as
ONEMA and AFB). The sampling protocols were standardized since 1995 and a
complete description of the sampling process can be found in
@danet_species_nodate (main text and appendixS2). The fishes are sampled using
electrofishing all across the streams in shallow streams and over the bank by
boat in deeper streams. All the fishes are counted and the bodysize of fishes are
recorded for a sample of them, enabling the estimate the distribution of
individual bodysize by species for each sampling [@bonnaffe_comparison_2021;
@danet_species_nodate].

We selected the sites that have been monitored at least 10 years on the
1995-2018 period, with the same sampling protocol over time. We selected
one sampling a year by site. In each site, we ensured low temporal heterogeneity
in sampling period by selecting sampling that were performed in the same period,
leading to a mean standard deviation of sampled month in each site of three
weeks [@danet_species_nodate]. Those steps leads to the selection of 
`r nb_st` sites (Figure \@ref(fig:matmet)A).

@@

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or three
life stages by fish species, each one being delimited by body size
[@bonnaffe_comparison_2021; @danet_species_nodate]. The ontogenic fish diet
database was filled according to fishbase (Froese et al. 2021) and literature
[@bonnaffe_comparison_2021; @danet_species_nodate].

@@

$$
\begin{aligned}
    Y_{i|n, t} &= \alpha + \betaTime_t + \epsilon_{i|n, t}
  (\#eq:gen)
\end{aligned}
$$

@@

The temporal trends at each site have been extracted from the hierarchical model
using mean BLUP values (Best Linear Unbiaised Predictor), then multiplied
by 10 to get temporal trends by decade. For logged transformed biomass and
species richness, we derived percentage change by decade by transforming the
slopes ($\beta\prime = (e^{\beta } - 1) \times 100 \times 10$).

@@

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were
standardizated and the confidence intervals were computed with semEff R package
[@murphy_semEff_2021]. Confidence intervals were obtained by bootstrapping the
linear models 1000 times and computing adjusted bootstrap percentile interval
(BCa), which are adjusted for asymetric distribution of resampling. The marginal
and conditional $R^2$ associated to the linear model of the SEMs were computed
the Nagelkerke method [@lefcheck_piecewisesem_2016].

@@

>>>>

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
#  bookdown::word_document2:
#    fontsize: 12pt
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
    fontsize: 12pt
    header-includes:
       - \usepackage{natbib}
       - \usepackage{float}
       - \usepackage{multirow}
       - \usepackage{booktabs}
       - \usepackage{makecell}
       - \usepackage{setspace}
       - \doublespacing
       - \usepackage{lineno}
       - \linenumbers
    geometry: margin=2cm
    number_sections: true
bibliography: references.bib
---

@@

```{r, results = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(viridis)
library(kableExtra)
library(sf)
mypath <- rprojroot::find_package_root_file
source(mypath("R/variable_shortcut.R"))
source(mypath("R/misc.R"))
source_dir(get_mypath("R")) # Define your custom code as a bunch of functions.

@@

```{r}
tar_load(full_data2)
nb_st <- unique(full_data2$station) %>% length
nb_op <- nrow(full_data2) %>% round(., -3)

rm(full_data2)
```

# Abstract (267 / 250 words max)

Understanding how biodiversity changes is a central question in ecology, but
still challenging because of the multidimensionality of biodiversity changes.
Although the structure of ecological networks is a crucial determinant of
community dynamics, this dimension of biodiversity changes have been greatly
overlooked. One main reason is the scarce availability of the temporal series of
ecological networks. Our knowledge of the relationships among structural
properties of ecological networks comes mainly from spatial comparison, assuming
a space-for-time equivalence. However, little is know about the validity of the
space-for-time hypothesis, i.e. if spatial comparison of ecological networks
accurately translates temporal sequences of assemblages. Here we questioned the
relationships between the temporal trends of food-web structure and the ones
of community biomass and species richness, and we assessed how those
relationships differed from the ones estimated in space. We inferred the
structure of using food diet and individual body size data from fish monitoring
in France. We estimated the temporal trends of food-web structure in 400 sites
using a hierarchical Bayesian model and then assess the relationships among them
using structural equation models. We found that temporal trends of connectance
and average trophic level are positively related to ones of community biomass
but surprisingly poorly related to the temporal trends of species richness. We
further found support for space-for-time hypothesis for community biomass, but
not for species richness. We suggest that our results resonates with the
energetic paradigm of food-webs. Our study sheds a new light on the dynamics of
food-web assemblage in natural settings and how the structure of ecological
networks changes according to the changes of other biodiversity facets.

# Significance (109 / 120 words max)

While the multidimensional nature of biodiversity changes has been clearly
emphasized, little is know how the structure of food-web changes according to
changes in species richness and community biomass. By inferring food-web
structure based on food-diet data and body size ratios, we analyzed how food-web
structure in time according to simultaneously changes in species richness and
community biomass. We identify that community biomass is the primary driver of
temporal changes in foodweb structure, but that species richness has surprisingly
low effect. We further show that space-for-time leads to contrasting results for
species richness. Our results shed a new light on how foodweb assemble
and disassemble in natural settings.

# Introduction (487 words)

There is rising recognition that biodiversity changes are
multidimensional[@mcgill_fifteen_2015; @blowes_geography_2019;
@hillebrand_biodiversity_2018; @blowes_local_nodate]. Studies found important
changes in species composition despite no changes in species richness in average
[@dornelas_assemblage_2014; @blowes_geography_2019], and highlighting the need
for further investigations on how biodiversity changes over time.

Yet species interactions are largely absent from biodiversity changes. Although
species interactions, described as ecological networks, drive for a large part
community composition, species coexistence, and ecosystem
functioning[@chesson_mechanisms_2000; @thebault_food-web_2003;
@oconnor_general_2017]. The structure of ecological networks can trigger or
impede perturbation propagations, i.e secondary extinctions[@paine_note_1969;
@dunne_network_2002; @srinivasan_response_2007;
@stouffer_compartmentalization_2011]. Higher connectance, which measures the
degree of generalism in a network, is predicted to increase robustness to
secondary extinctions[@dunne_network_2002; @sole_complexity_2001] and studies
found that higher trophic level was associated to a lower temporal variability
of community biomass in theoretical models and empirical
data[@shanafelt_stability_2018; @danet_species_2021]. Then, accounting for how
the structure of ecological networks changes over time according to other
biodiversity facets is crucial.

Th availability of ecological network timeseries is
scarce[@kortsch_disentangling_2021], such as predictions about how the structure
of ecological networks change according to changes in species richness come from
spatial comparison of networks. Such studies showed that connectance is
negatively related to species richness in small (i.e. below 20 species)
empirical food-webs [reviewed in @dunne_network_2006], which reflects
constraints on the number of realizable species interactions or an increasing
proportion of specialists in more rich foodwebs. Connectance has been shown to
scale positively or negatively with community biomass in theoretical
food-webs[@thebault_food-web_2003; @poisot_trophic_2013] while the average
trophic level was previously shown to scale negatively with species
richness[@danet_species_2021].

@@

The increasing availability ecological temporal series and the development of
methods to reconstruct food-webs[@bonnaffe_comparison_2021;
@gravel_inferring_2013; @danet_species_2021] makes possible to overcome those
research gaps. We used a standardized monitoring of fish riverine communities on
the period 1995-2018 over France, documenting the individual size of fishes in
`r nb_st` sites during ten years or more. We reconstructed around `r nb_op`
food-webs based on allometric relationships and food diet data, and then
computed connectance, and average trophic level as well as community biomass and
species richness. We assessed how the local structure of food-webs changes over
time according to simultaneous changes in community biomass and species richness
over time and then if the relationships between changes in food-web structure,
community biomass and species richness, are the same over space and over time,
i.e. is the space-for-time substitution valid?

@@

Using a hierarchical model to estimate conjointly the temporal trends of  
foodweb structure at all sites, we found weak evidence of a global decrease in
community biomass (CI80%:`r gbm` by decade; median [Credible Interval]) but of a
global increase in species richness (CI80%: `r gr` by decade) across the `r nb_st` sites. 
We did not find evidence of a global trends in connectance and average trophic
level (CI80%: `r gct` and `r gt` respectively). The hierarchical model captured
well the variance associated to the spatial structure of the data with
conditional $R^2$ going from  `r r2ic$log_bm_std` (mean [CI95%]) for community
biomass to `r r2ic$log_rich_std` for species richness (Table S1).

```{r}
sem_cap <- "(a) Structural equation models linking the temporal trends of the food-web
structure. The path widths are proportional to the standardized slope
coefficients. The standardized slope coefficients and their associated 95%
confidence intervals are displayed along the path. The marginal R-squared of the
linear models are displayed. (b to f) Direct bivariate relationships derived
from the structural equation model. The distribution of the site temporal trends
are displayed in Fig. S1."
```

@@

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r rdtps$ct$bm`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that decreased
in biomass over time lose highly connected trophic species over time, and
reversely that communities gaining biomass over time also gain highly connected
trophic species over time. We found a negative relationship between the temporal
trends of connectance and the ones of species richness
($r_\delta = `r rdtps$ct$r`$, Fig. \@ref(fig:sem)d). We further found the
species richness had an indirect positive effect on connectance, through
community biomass ($r_\delta = `r ridtps$ct$r`$). It resulted in a relatively
weak total negative effect of species richness on connectance
($r_\delta = `r rttps$ct$r`$), that we also found in random extinction
simulation (Fig. S2).

@@

We found that species richness scaled positively with community biomass, both in
spatial and temporal analysis. This result is in accordance with previous
analysis across space in stream communities [@woods_testing_2020;
@danet_species_2021] but also in grasslands, salt marches and forests
[@hector_plant_1999; @tilman_biodiversity_2014; @wu_relationship_2015;
@li_relationship_2018; but see @grace_integrative_2016]. Therefore, our results
suggest that Biodiversity-Ecosystem functioning prediction also hold in the
temporal dynamics of food-webs. The positive relationship between species
richness and total biomass might be due to positive complementarity in resource
use [@carey_determining_2011], which is held by niche differentiation
[@loreau_biodiversity_1998; @loreau_partitioning_2001] or to the higher
probability of getting high biomass species with higher species
richness [@loreau_biodiversity_2001].

We found that both connectance and average trophic level had a positive
relationship with community biomass over time, meaning that connectance and
average trophic level increase over time when community biomass increases over
time as well, and likewise that connectance and average trophic level decrease
over time when community biomass decreases over time. We also found that the
temporal trends of connectance and average trophic level were strongly
correlated (Fig. S2). All together, it suggests that the disturbances that
lead to decline or increase of community biomass over time are concomitant to
the decline and increase respectively of trophic species at high trophic level.
Likewise, we found that the temporal trends of community biomass were positively
associated to the temporal trends of both number and proportion of piscivorous
fish species present in the communities (Fig. S2). This result is in
accordance with the energetic paradigm of food-webs which proposes that
food-webs are primarily constrained by fluxes of matter and energy
[@lindeman_trophic-dynamic_1942], which suggest that productivity constraints
food chain length by energy limitation [@jenkins_productivity_1992;
@takimoto_ecosystem_2008]. Higher trophic levels have been shown to be more
sensible to ecosystem size [@takimoto_ecosystem_2008; @post_ecosystem_2000], to
resource availability and disturbance in small systems [@post_long_2002] and
particularly to disturbance in hydrologic regimes in streams [@sabo_role_2010;
@mchugh_dual_2010; @jake_vander_zanden_global_2007]. In the same line, the
metabolic theory of ecology also predicts that top predators are highly
constrained by energy limitation [@arim_relationship_2007]. However this
interpretation does hold if the temporal trends of community biomass are related
to the quantity of resource available. Another explanation can be that top
trophic level species are often the ones that are targeted, harvested and those
that are artificially introduced for recreational fisheries
[@reid_emerging_2019]. Further investigations on the drivers of community
biomass decline and increase over time would be needed to further conclude.

In contrast, we found that the temporal trends of connectance was negatively
related to the temporal trends of species richness, in accordance with the
strong negative relationships that we found in the present study and from
previous results over spatial gradient [@dunne_network_2006;
@danet_species_2021]. Furthermore, we found that temporal trends of the
proportion of piscivorous species was negatively associated with the temporal
trends of species richness (Fig. S2), meaning that communities that increase in
richness over time have an increase in the proportion of non-piscivore over time
and vice versa. This result is in accordance with a meta-analysis showing the
number of species by trophic level decrease toward higher trophic level
[@turney_pyramids_2016]. It suggests that the temporal increase of species
richness found in our dataset is associated with a decrease of connectance
because species richness increases are biased toward non-piscivorous species
which have no piscivory links. We also found that the changes of average trophic
level over time were not related to the changes of species richness over time.
All together, an explanation can be that species richness changes over time,
without changes in community biomass, are more linked to non piscivorous species
having lower trophic level and relatively low biomass, which results in negative
relationships with temporal trends in connectance and no relationship with
average trophic level.

However, as we found a strong positive relationship between the temporal trends
of species richness and community biomass, it results that species richness
temporal trends had an indirect positive effect on the temporal trends of both
connectance and average trophic level via community biomass. Then, our study
suggests that species richness and community biomass have antagonistic effects
on connectance, respectively negative and positive. The temporal trends of
species richness had then a total effect on the temporal trends of connectance
and average trophic level which was twice lower than the one of the temporal
trends of community biomass. It suggests that the temporal assembly and
disassembly of food-webs are more sensitive to variations in community biomass
than in species richness, in other words that food-web structure is more robust
to changes in species richness than in biomass over time. Our study then
suggests that the changes of food-web structure over time related to changes in
species richness could be actually mediated by changes in community biomass.

As studies reported community biomass or species richness temporal trends
[@hallmann_more_2017; @ceballos_biological_2017; @dornelas_assemblage_2014;
@blowes_geography_2019], we document that temporal trends of species richness
and community biomass might have different consequences on the structure of the
community. In particular, the decrease of the average trophic level associated
to the loss of community biomass could have cascading effect on ecosystem
functioning as the highest predators are predicted to increase productivity and
stabilize communities [@wang_biodiversity_2018; @shanafelt_stability_2018;
@danet_species_2021]. In contrast, the loss of species richness might be linked
to loss of redundancy in the food-web, highlighting that those communities may
become more at risk regarding future perturbations [@yachi_biodiversity_1999].
Interestingly, the hypothesis that temporal trends in community biomass is
related to higher trophic level in our system has a strong reasoning with the
effect of human fisheries in marine systems which affect primarily fish species
at top trophic levels [@sprules_surfing_2016].

Regarding the space-for-time substitution, we found that connectance was
positively linked to community biomass both over space and time. This result
resonates with theoretical work showing that higher food-web connectance can
result in higher community biomass [@thebault_food-web_2003], or lower
[@poisot_trophic_2013] over space, highlighting that the relation between
connectance and community biomass may be dependant on how food-web structure
change with community biomass over space and time. Likewise, we found that the
average trophic level had a positive relationship with community biomass over
space and time, in accordance with a theoretical study over space predicting an
increase of community biomass subsequent to the presence of higher predators
[@wang_biodiversity_2018]. Finally, we found a positive relationship between
community biomass and species richness both in space and time, however the
temporal relationship was stronger. We also found a negative relationship
between connectance and species richness space and time, in accordance with
previous findings over space in empirical food-webs [@dunne_network_2006]. We further
found no relationship between average trophic level and species richness over
time while we found a twice stronger but positive relationship between median
average trophic level and median species richness over the spatial gradient. An
explanation might be that environmental gradient and the pattern of fish
migration since the last glaciation, for the strong negative effect of species
richness on both connectance and average trophic level over spatial gradient.
However, we controlled our spatial analysis by the Biological Demand in Oxygen,
water temperature and the stream gradient. This result is also in line with
other studies which found a mismatch between spatial and temporal patterns
[@adler_contrasting_2007; @johnson_testing_2008]. This discrepancy in space
and time might be due to the strong assumption that variations in space reflect
differences in assembly stages of ecological communities and then mirror their
temporal assembly/disassembly trajectories. Temporal variation of species
richness might relate to transitory dynamics while spatial gradients reflect
aggregated dynamics over a longer time scale including a far number of processes
including high species turnover along spatial gradients.

@@

Our study suggests that species richness and community biomass temporal trends
in natural fish communities relate on different ecological processes, biomass
temporal dynamics being related to variations of food-web structure at the
higher trophic levels while variations of species richness is weakly related to
change in food-web structure. Our result also suggests that space for time
substitution may be relevant to predict the effects of community biomass losses
or gains on food-web structure but also that space-for-time substitution might
not work for species richness. Our study calls for further investigations to
support the generality of assembly rules in food-webs. They are crucial to
better understand the assembly and disassembly of food-webs and then predict its
impact on ecosystem functioning.

@@

We used a monitoring of freshwater fishes conducted all over France and
conducted by the Office Français de la Biodiversité (OFB, previously known as
ONEMA and AFB). The sampling protocols were standardized since 1995 and a
complete description of the sampling process can be found in
@danet_species_2021 (main text and appendix S2). The fishes are sampled using
electrofishing all across the streams in shallow streams and over the bank by
boat in deeper streams. All the fishes are counted and the bodysize of fishes are
recorded for a sample of them, enabling the estimate the distribution of
individual bodysize by species for each sampling [@bonnaffe_comparison_2021;
@danet_species_2021].

We selected the sites that have been monitored at least 10 years on the
1995-2018 period, with the same sampling protocol over time. We selected
one sampling a year by site. In each site, we ensured low temporal heterogeneity
in sampling period by selecting sampling that were performed in the same period,
leading to a mean standard deviation of sampled month in each site of three
weeks [@danet_species_2021]. Those steps leads to the selection of 
`r nb_st` sites (Figure \@ref(fig:matmet)A).

@@

We determined the feeding interactions of trophic species according to their
species identity and their body size. The diet database described two or three
life stages by fish species, each one being delimited by body size
[@bonnaffe_comparison_2021; @danet_species_2021]. The ontogenic fish diet
database was filled according to fishbase (Froese et al. 2021) and literature
[@bonnaffe_comparison_2021; @danet_species_2021].

@@

$$
\begin{aligned}
    Y_{i|n, t} &= \alpha + \beta Time_t + \epsilon_{i|n, t}
  (\#eq:gen)
\end{aligned}
$$

@@

The temporal trends at each site have been extracted from the hierarchical model
using mean BLUP values (Best Linear Unbiaised Prediction), then multiplied
by 10 to get temporal trends by decade. For logged transformed biomass and
species richness, we derived percentage change by decade by transforming the
slopes ($\beta\prime = (e^{\beta } - 1) \times 100 \times 10$).

@@

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were
standardizated and the confidence intervals were computed with semEff R package
[@murphy_semEff_2021]. Confidence intervals were obtained by bootstrapping the
linear models 100 times and computing adjusted bootstrap percentile interval
(BCa), which are adjusted for asymetric distribution of resampling. The marginal
and conditional $R^2$ associated to the linear model of the SEMs were computed
with the Nagelkerke method [@lefcheck_piecewisesem_2016].

@@

# Extinction simulation in the metaweb

We simulated extinction sequences in the metaweb according to three scenarios to get expectations on how the food-web structure should change according to the addition and deletion of species and nodes. The scenario of extinction sequences were related to the trophic level of the species/nodes. In the decreasing and increasing scenario, we removed species/nodes in decreasing, resp. increasing order, of their trophic level, i.e. respectively the higest and lowest trophic level first. In the random extinction sequence, we built 50 random extinction sequences. For each simulation, we removed fish species/nodes that lost all their preys. We computed the connectance, the average and the maximum trophic level.
